Lesson 3
========================================================

***

### What to Do First?
Notes: 

***

### Pseudo-Facebook User Data
Notes:

```{r Pseudo-Facebook User Data}
getwd()
#setwd('/Users/pure/Private_Local_Data/Study/Udacity/DataAnalyst/Lessons/4.\ Exploratory\ Data\ Analysis/eda-course-materials/lesson3/')
list.files()
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')
```

***

### Histogram of Users' Birthdays
### http://flowingdata.com/2014/02/27/how-to-read-histograms-and-use-them-in-r/
Notes:
```{r}
#install.packages('ggplot2')
library(ggplot2)
```

```{r Histogram of Users\' Birthdays}
names(pf)

qplot(x = dob_day, data = pf) +
  scale_x_continuous(breaks=1:31)
```
```{r}
ggplot(aes(x = dob_day), data = pf) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = 1:31)
```

***

#### What are some things that you notice about this histogram?
Response:

***

### Moira's Investigation
Notes:

***

### Estimating Your Audience Size
Notes:

***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response:

#### How many of your friends do you think saw that post?
Response:

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response:

***

### Perceived Audience Size
Notes:

***
### Faceting
### http://www.cookbook-r.com/Graphs/Facets_(ggplot2)/
Notes:

```{r Faceting}
qplot(x = dob_day, data = pf) +
  scale_x_continuous(breaks=1:31)+
  facet_wrap(~dob_month, ncol = 3)
```
```{r}
ggplot(aes(x = dob_day), data = pf) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = 1:31) +
  facet_wrap(~dob_month)
```

#### Let’s take another look at our plot. What stands out to you here?
Response:

***

### Be Skeptical - Outliers and Anomalies
Notes:
A top-coded (http://en.wikipedia.org/wiki/Top-coded) data set is one for which the value of variables above an upper bound are censored. This is often done to preserve the anonymity of people participating in the survey.
***

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira’s outlier?
Response:

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
qplot(x = friend_count, data = pf)
```
```{r}
ggplot(aes(x = friend_count), data = pf) +
  geom_histogram()
```

#### How is this plot similar to Moira's first plot?
Response:

***

### Limiting the Axes
Notes:

```{r Limiting the Axes}
qplot(x = friend_count, data = pf, xlim = c(0, 1000))
```

```{r}
qplot(x = friend_count, data = pf) +
  scale_x_continuous(limits = c(0, 1000))
```

```{r}
ggplot(aes(x = friend_count), data = pf) +
  geom_histogram() +
  scale_x_continuous(limits = c(0, 1000))
```



### Exploring with Bin Width
Notes:

***

### Adjusting the Bin Width
### http://docs.ggplot2.org/current/scale_continuous.html
Notes:
```{r}
qplot(x = friend_count, data = pf, binwidth = 25) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50))
```

```{r}
ggplot(aes(x = friend_count), data = pf) +
  geom_histogram(binwidth = 25) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 100))
```

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = pf, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```

```{r}
qplot(x = friend_count, data = pf) +
  facet_grid(gender ~ .)
```

```{r}
ggplot(aes(x = friend_count), data = pf) +
  geom_histogram() +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```

***

### Omitting NA Values
Notes:

```{r Omitting NA Values}
qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```

```{r}
qplot(x = friend_count, data = na.omit(pf), binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```

```{r}
ggplot(aes(x = friend_count), data = subset(pf, !is.na(gender))) +
  geom_histogram() +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```



***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}
table(pf$gender)
by(pf$friend_count, pf$gender, summary)
```

#### Who on average has more friends: men or women?
Response:

#### What's the difference between the median friend count for women and men?
Response:

#### Why would the median be a better measure than the mean?
Response:
Median is always half of the data.
***

### Tenure
### colour hex code
### ggplot theme: http://ggplot2.tidyverse.org/reference/theme.html

Notes:

```{r Tenure}
qplot(x = tenure, data = pf, binwidth = 30,
      color = I('black'), fill = I('#099DD9'))
```

```{r Tenure1}
ggplot(aes(x = tenure), data = pf) +
  geom_histogram(binwidth = 30, color = 'black', fill = '#099DD9')
```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}
qplot(x = tenure/365, data = pf, binwidth = .25,
  color = I('black'), fill = I('#F79420')) +
  scale_x_continuous(breaks = seq(1, 7, 1), limits = c(0, 7))
```

```{r Tenure3}
ggplot(aes(x = tenure/365), data = pf) +
  geom_histogram(binwidth = .25, color = 'black', fill = '#F79420')
```

***

### Labeling Plots
Notes:

```{r Labeling Plots}
qplot(x = tenure/365, data = pf, binwidth = .25,
  color = I('black'), fill = I('#F79420')) +
  scale_x_continuous(breaks = seq(1, 7, 1), limits = c(0, 7)) +
  xlab('Number of years using Facebook') +
  ylab('Number of users in sample')
```

```{r}
ggplot(aes(x = pf$tenure/365), data = pf) +
  geom_histogram(binwidth = .25, color = 'black', fill= '#F79420') +
  scale_x_continuous(breaks = seq(1, 7, 1), limits = c(0, 7)) +
  xlab('Number of years using Facebook') +
  ylab('Number of users in sample')
```


***

### User Ages
Notes:

```{r}
qplot(x = pf$age, data=pf, binwidth = 1, 
      color = I('black'), fill = I('#5760AB')) + 
  scale_x_continuous(breaks = seq(0, 113, 5))
```


```{r User Ages}
ggplot(aes(x = pf$age), data = pf) +
  geom_histogram(binwidth = 1, color = 'black', fill= '#5760AB') +
  scale_x_continuous(breaks = seq(0, 113, 5))
```


#### What do you notice?
Response:

***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes:

***

### Transforming Data
Notes:
```{r}
p1 <- ggplot(aes(x = friend_count), data = pf) +
  geom_histogram(color = 'black', fill='#5760AB')

p2 <- qplot(x = log10(pf$friend_count+1), data = pf, color = I('black'), fill = I('#5760AB'))
p3 <- p1 + scale_x_log10()

p4 <- p1 + scale_x_sqrt()

library(gridExtra)
grid.arrange(p1, p2, p3, p4, ncol = 2)
```

***

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}
# Done in above
```

***


### Frequency Polygons

```{r Frequency Polygons1}
qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth=10, geom = 'freqpoly', color = gender) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50))
  #facet_wrap(~gender)
```
```{r Frequency Polygons2}
qplot(x = friend_count, y = ..count../sum(..count..), data = subset(pf, !is.na(gender)), binwidth=10, geom = 'freqpoly', color = gender) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  xlab('Friend Count') +
  ylab('Proportion of users with that friend count')
  #facet_wrap(~gender)
```

```{r}
ggplot(aes(x = friend_count, y = ..count../sum(..count..)),
       data = subset(pf, !is.na(gender))) +
  geom_freqpoly(aes(color = gender), binwidth=10) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  xlab('Friend Count') +
  ylab('Proportion of users with that friend count')
```

***

### Likes on the Web
Notes:
```{r}
qplot(x = www_likes, data = subset(pf, !is.na(gender)), 
      geom = 'freqpoly', color = gender) +
  scale_x_log10()
```

```{r Likes on the Web1}
ggplot(aes(x = www_likes), data = subset(pf, !is.na(gender))) +
  geom_freqpoly(aes(color = gender)) +
  scale_x_log10()
```
```{r Likes on the Web2}
ggplot(aes(x = www_likes, y=..density..), data = subset(pf, !is.na(gender))) +
  #geom_histogram(binwidth = 10, color = 'black', fill = '#5760AB') +
  geom_freqpoly(aes(color=gender)) +
  scale_x_log10()
  #scale_x_continuous(limits = c(0, 100), breaks = seq(0, 1000, 100))
  #facet_wrap(~gender)
```

```{r}
by(pf$www_likes, pf$gender, sum)
```


***

### Box Plots
Notes:

```{r Box Plots}
qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 25) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.
ylim: When we use the ylim parameter, we're actually removing data points from our calculations. So the original box plots that we have might look slightly different. 

scale_y_continuous: Also this value might not be accurate for all of our data. When we use the ylim parameter of the scale_y_continuous layer, we actually remove data points from calculations.

coord_cartesian: Better way. 
```{r}
p1 <- qplot(x = gender, y = friend_count, 
      data = subset(pf, !is.na(gender)), 
      geom = 'boxplot',
      ylim = c(0, 1000))

p2 <- qplot(x = gender, y = friend_count, 
      data = subset(pf, !is.na(gender)), 
      geom = 'boxplot') +
  scale_y_continuous(limits = c(0, 1000))

p3 <- qplot(x = gender, y = friend_count,
      data = subset(pf, !is.na(gender)),
      geom = 'boxplot') + 
  coord_cartesian(ylim = c(0, 1000))

grid.arrange(p1, p2, p3, nrow = 1)
```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}
qplot(x = gender, y = friend_count,
      data = subset(pf, !is.na(gender)),
      geom = 'boxplot') + 
  coord_cartesian(ylim = c(0, 250))
```

```{r}
by(pf$friend_count, pf$gender, summary)
```


#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}
qplot(x = gender, y = friendships_initiated, 
      data = subset(pf, !is.na(gender)),
      geom = 'boxplot') +
  coord_cartesian(ylim = c(0, 150))
```

```{r}
by(pf$friendships_initiated, pf$gender, summary)
```

Response:

***

### Getting Logical
Notes:

```{r Getting Logical}
summary(pf$mobile_likes)
```

```{r}
summary(pf$mobile_likes > 0)
```

```{r}
pf$mobile_check_in <- NA
pf$mobile_check_in <- ifelse(pf$mobile_likes > 0, 1, 0)
pf$mobile_check_in <- factor(pf$mobile_check_in)
summary(pf$mobile_check_in)
```

```{r}
sum(pf$mobile_check_in == 1) / length(pf$mobile_check_in) * 100
```


Response:
65% of users have mobile experience.
***

### Analyzing One Variable
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!