Lesson 5
========================================================

### Multivariate Data
Notes:
```{r}
getwd()
#setwd('/Users/pure/Private_Local_Data/Study/Udacity/DataAnalyst/Lessons/4.\ Exploratory\ Data\ Analysis/eda-course-materials/lesson3/')
setwd('/media/pure/SharedData/Development/Personal_Dev/DataAnalyst/Lessons/4. Exploratory Data Analysis/eda-course-materials/lesson3/')
list.files()

library(ggplot2)
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')
### You can also read in the data using the following code:
#pf < - read.delim('pseudo_facebook.tsv')
```
***

### Moira Perceived Audience Size Colored by Age
Notes:
Moira's Investigation from Lesson 3 
(https://classroom.udacity.com/courses/ud651/lessons/755618712/concepts/8140986010923)

Bernstein, M. S., Bakshy, E., Burke, M., & Karrer, B. (2013). Quantifying the invisible audience in social networks. In Proceedings of the SIGCHI Conference on Human Factors in Computing Systems (CHI 2013), pp. 21-30. 
(http://hci.stanford.edu/publications/2013/invisibleaudience/invisibleaudience.pdf)
***

### Third Qualitative Variable
Notes:

Important Notice! Please note that in newer versions of dplyr (0.3.x+), the syntax %.% has been deprecated and replaced with %>%. Videos in the course use %.%, which will produce warning messages. If you answer using the chain operator, you should use %>% instead.

Another warning: Version 0.4.0 of dplyr has a bug when using the median function on the summarize layer, depending on the nature of the data being summarized. You may need to cast the data as a numeric (float) type to get the expected results, e.g. median(as.numeric(var)).

You can include multiple variables to split the data frame when using group_by() function in the dplyr package.

new_groupings <- group_by(data, variable1, variable2)

Repeated use of summarise() and group_by(): (http://stackoverflow.com/questions/21653295/dplyr-issues-with-group-by) The summarize function will automatically remove one level of grouping (the last group it collapsed).


```{r}
library(ggplot2)
suppressMessages(library(dplyr))
```

```{r}
ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 4)
```

```{r}
ggplot(aes(x = age, y = friend_count), 
       data = subset(pf, !is.na(gender))) +
  geom_line(aes(color = gender), stat = 'summary', fun.y = median)
```

```{r Third Qualitative Variable}
pf.fc_by_age_gender <- pf %>%
  filter(!is.na(gender)) %>%
  group_by(age, gender) %>%
  summarise(mean_friend_count = mean(friend_count),
            median_friend_count = median(friend_count),
            n = n()) %>%
  ungroup() %>%
  arrange(age)
```
Write code to create a new data frame, called 'pf.fc_by_age_gender', that contains
information on each age AND gender group.

The data frame should contain the following variables:

  mean_friend_count,
  median_friend_count,
  n (the number of users in each age and gender grouping)

Here is an example of the structure of your data frame. Your
data values will be different. Note that if you are grouping by
more than one variable, you will probably need to call the
ungroup() function. 

  age gender mean_friend_count median_friend_count    n
1  13 female          247.2953                 150  207
2  13   male          184.2342                  61  265
3  14 female          329.1938                 245  834
4  14   male          157.1204                  88 1201

***

### Plotting Conditional Summaries
Notes:
Create a line graph showing the median friend count over the ages
for each gender. Be sure to use the data frame you just created,
pf.fc_by_age_gender.

Your code should look similar to the code we used to make the plot the first time. 
It will not need to make use of the stat and fun.y parameters.

ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(gender))) +
  geom_line(aes(color = gender), stat = 'summary', fun.y = median)
  
You could also use color = gender inside the aes() wrapper of ggplot().


```{r Plotting Conditional Summaries}
ggplot(aes(x = age, y = median_friend_count),
       data = pf.fc_by_age_gender) +
  geom_line(aes(color = gender))
```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:
You can also restructure the data using the tidyr package. You can review examples and how to use the package in the Data Wrangling with R (https://s3.amazonaws.com/udacity-hosted-downloads/ud651/DataWranglingWithR.pdf) pdf.

The code to change the data frame from long format to wide (or tidy format) is shown below. I encourage you to read the Data Wrangling pdf and write code using the tidyr package before looking at the solution below. 

install.packages("tidyr")
library(tidyr)

spread(subset(pf.fc_by_age_gender, 
       select = c('gender', 'age', 'median_friend_count')), 
       gender, median_friend_count)
       
I think you will find the tidyr package easier to use than the reshape2 package. Both packages can get the job done.

An Introduction to reshape2 by Sean Anderson (http://www.seananderson.ca/2013/10/19/reshape.html)

Converting Between Long and Wide Format (http://www.cookbook-r.com/Manipulating_data/Converting_data_between_wide_and_long_format/)

Melt Data Frames (http://www.r-bloggers.com/melt/)
***
```{r}
#install.packages("tidyr")
library(tidyr)
```
```{r}
spread(subset(pf.fc_by_age_gender, 
              select = c('gender', 'age', 'median_friend_count')),
       gender, median_friend_count)
```


### Reshaping Data
Notes:
It’s important to use quotes around the variable name that is assigned tovalue.var. 

We could also create a similar data frame using the dplyr and tidyr packages:

```{r reshape}
#install.packages('reshape2')
library(reshape2)

pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender,
                                  age ~ gender,
                                  value.var = 'median_friend_count')

head(pf.fc_by_age_gender.wide)
```


```{r Alternative method for reshaping}
pf.fc_by_age_gender.wide <-
  subset(pf.fc_by_age_gender[c('age', 'gender', 'median_friend_count')],
         !is.na(gender)) %>%
  spread(gender, median_friend_count) %>%
  mutate(ratio = male / female)

head(pf.fc_by_age_gender.wide)
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
ggplot(aes(x = age, y = male/female), data = pf.fc_by_age_gender.wide) +
  geom_line() +
  geom_hline(yintercept = 1, linetype = 'dashed', color = 'blue')
```

***

### Third Quantitative Variable
Notes:
Create a variable called year_joined in the pf data frame using the variable tenure and 2014 as the reference year.

The variable year_joined should contain the year that a user joined facebook.

See the Instructor Notes for three hints if you get stuck. Scroll down slowly to see one hint at a time if you would like some guidance.

A common mistake is to use tenure rather than pf$tenure or with(pf, tenure...). Remember that you need to access the variable in the data frame. This is not one of the hints! :) 

Hint 1: Divide the tenure variable by a number. Tenure is measured in days, but we want to convert it to years. 

Hint 2: Subtract tenure measured in years from 2014. What does the decimal portion represent? Should we round up or round down to the closest year? 

Hint 3: You can use the floor() function to round down to the nearest integer. You can use the ceiling() function to round up to the nearest integer. Which one should you use?

```{r Third Quantitative Variable}
pf$year_joined <- floor(2014 - pf$tenure/365)
```

***

### Cut a Variable
Notes:

Create a new variable in the data frame called year_joined.bucket by using the cut function on the variable year_joined.

You need to create the following buckets for the new variable, year_joined.bucket

      (2004, 2009]
      (2009, 2011]
      (2011, 2012]
      (2012, 2014]

Note that a parenthesis means exclude the year and a bracket means include the year.

Look up the documentation for cut or try the link in the Instructor Notes to accomplish this task.

The Cut Function (http://www.r-bloggers.com/r-function-of-the-day-cut-2/)

Be sure to close your parentheses!

A common mistake is to use year_joined rather than pf$year_joined or with(pf, year_joined...). Remember that you need to access the variable in the data frame.

```{r}
summary(pf$year_joined)
```
```{r}
table(pf$year_joined)
```


```{r Cut a Variable}
pf$year_joined.bucket <- cut(pf$year_joined, breaks = c(2004, 2009, 2011, 2012, 2014))
```

***

### Plotting it All Together
Notes:
Create a line graph of friend_count vs. age so that each year_joined.bucket is a line tracking the median user friend_count across age. This means you should have four different lines on your plot.

You should subset the data to exclude the users whose year_joined.bucket is NA.
```{r}
table(pf$year_joined.bucket, useNA = 'ifany')
```
```{r Example Code}
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(gender))) +
  geom_line(aes(color = gender), stat = 'summary', fun.y = median)
```
```{r}
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(gender))) +
  geom_line(aes(color = gender), stat = 'summary', fun.y = median)
```

```{r Plotting it All Together}
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(year_joined.bucket))) +
  geom_line(aes(color = year_joined.bucket), 
            stat = "summary", 
            fun.y = median)
```

***

### Plot the Grand Mean
Notes:
Write code to do the following:

(1) Add another geom_line to code below to plot the grand mean of the friend count vs age.

(2) Exclude any users whose year_joined.bucket is NA.

(3) Use a different line type for the grand mean.

As a reminder, the parameter linetype can take the values 0-6:

0 = blank, 1 = solid, 2 = dashed
3 = dotted, 4 = dotdash, 5 = longdash
6 = twodash
```{r Plot the Grand Mean}
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(year_joined.bucket))) +
  geom_line(aes(color = year_joined.bucket), 
            stat = "summary", 
            fun.y = mean) +
  geom_line(stat = "summary", fun.y = mean, linetype = 2)
```

***

### Friending Rate
Notes:

What is the median friend rate?
.2205
What is the maximum friend rate?
417

```{r Friending Rate}
with(subset(pf, tenure >= 1), summary(friend_count/tenure))
```

***

### Friendships Initiated
Notes:

Understanding the Bias-Variance Tradeoff (http://scott.fortmann-roe.com/docs/BiasVariance.html)

NOTE: The code changing the binning is substituting x = tenure in the plotting expressions with x = 7 * round(tenure / 7), etc., binning values by the denominator in the round function and then transforming back to the natural scale with the constant in front.

Create a line graph of mean of friendships_initiated per day (of tenure) vs. tenure colored by year_joined.bucket.

You need to make use of the variables tenure, friendships_initiated, and year_joined.bucket.

You also need to subset the data to only consider user with at least one day of tenure.

```{r Friendships Initiated}
ggplot(aes(x = tenure, y = friendships_initiated/tenure), data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket))
```

***

### Bias-Variance Tradeoff Revisited
Notes:

Instead of geom_line(), use geom_smooth() to add a smoother to the plot. You can use the defaults for geom_smooth() but do color the line by year_joined.bucket

```{r}
library(gridExtra)

p1 <- ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket), 
              stat = 'summary',
              fun.y = mean)

p2 <- ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket), 
              stat = 'summary',
              fun.y = mean)

p3 <- ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket), 
              stat = 'summary',
              fun.y = mean)

p4 <- ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket), 
              stat = 'summary',
              fun.y = mean)

grid.arrange(p1, p2, p3, p4, nrow = 4)
```
```{r Bias-Variance Tradeoff Revisited}
ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_smooth(aes(color = year_joined.bucket))
```


***

### Sean's NFL Fan Sentiment Study
Notes:
The Emotional Highs and Lows of the NFL Season 
(https://www.facebook.com/notes/facebook-data-science/the-emotional-highs-and-lows-of-the-nfl-season/10152033221418859)

The National Football League (NFL) 
(http://en.wikipedia.org/wiki/National_Football_League)

Sean Taylor
(http://seanjtaylor.com/)
***

### Introducing the Yogurt Data Set
Notes:
Download the yogurt data set (https://s3.amazonaws.com/udacity-hosted-downloads/ud651/yogurt.csv).

Bayesian Statistics and Marketing (http://www.perossi.org/home/bsm-1) contains the data set and a case study on it.

The citation for the original paper on the yogurt data set is Kim, Jaehwan, Greg M. Allenby, and Peter E. Rossi. "Modeling consumer demand for variety." Marketing Science 21.3 (2002): 229-250.

A special thanks to Professor Allenby for helping us understand this data set.

To learn more about scanner data, check out Panel Data Discrete Choice Models of Consumer Demand (http://www.nuff.ox.ac.uk/Economics/papers/2013/Panel%20Data%20Demand%20-%20June%2020,%202013.pdf)
***

### Histograms Revisited
Notes:

```{r}
yo <- read.csv("yogurt.csv")
str(yo)
```
```{r}
# change the id from an int to a factor
yo$id <- factor(yo$id)
str(yo)
```

```{r Histograms Revisited}
ggplot(aes(x = price), data = yo) +
  geom_histogram()

# very biased model
ggplot(aes(x = price), data = yo) +
  geom_histogram(binwidth = 10)
```

***

### Number of Purchases
Notes:
Create a new variable called all.purchases, which gives the total counts of yogurt for each observation or household.

One way to do this is using the transform function. You can look up the function transform and run the examples of code at the bottom of the documentation to figure out what it does.

The transform function produces a data frame so if you use it then save the result to 'yo'!

OR you can figure out another way to create the variable.

```{r}
summary(yo)
```

```{r}
length(unique(yo$price))
```

```{r}
table(yo$price)
```

The transform function takes in a data frame and allows us to add different variables to it by recombining variables that are already within the data frame.
```{r Number of Purchases}
yo <- transform(yo, all.purchases = strawberry + blueberry + pina.colada + plain + mixed.berry)
```

```{r Alternative vervose code}
#yo$all.purchases <- yo$strawberry + yo$blueberry + yo$pina.colada + yo$plain + yo$mixed.berry
```

***

### Prices over Time
Notes:
# Create a scatterplot of price vs time.

# This will be an example of a time series plot.

# Resolve overplotting issues by using
# techniques you learned in Lesson 4.

```{r Prices over Time}
ggplot(aes(x = time, y = price), data = yo) +
  geom_jitter(alpha = 1/4, shape = 21, fill = I('#F79420'))
```
***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

Note: x %in% y returns a logical (boolean) vector the same length as x that says whether each entry in x appears in y. That is, for each entry in x, it checks to see whether it is in y. This allows us to subset the data so we get all the purchases occasions for the households in the sample. Then, we create scatterplots of price vs. time and facet by the sample id. 


Use the pch or shape parameter to specify the symbol when plotting points. Scroll down to 'Plotting Points' on QuickR's Graphical Parameters.

```{r Looking at Sample of Households}
set.seed(4230)
sample.ids <- sample(levels(yo$id), 16)

ggplot(aes(x = time, y = price),
       data = subset(yo, id %in% sample.ids)) +
  facet_wrap( ~ id) +
  geom_line() +
  geom_point(aes(size = all.purchases), pch = 1)
```
```{r}
set.seed(10)
sample.ids <- sample(levels(yo$id), 16)

ggplot(aes(x = time, y = price),
       data = subset(yo, id %in% sample.ids)) +
  facet_wrap( ~ id) +
  geom_line() +
  geom_point(aes(size = all.purchases), pch = 1)
```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:
Here's the scatterplot matrix as a pdf (https://s3.amazonaws.com/udacity-hosted-downloads/ud651/scatterplotMatrix.pdf).

You'll need to run the code install.packages('GGally') to install the package for creating this particular scatterplot matrix.

If the plot takes a long time to render or if you want to see some of the scatterplot matrix, then only examine a smaller number of variables. You can use the following code or select fewer variables. We recommend including gender (the 6th variable)!

pf_subset = pf[, c('age', 'dob_year', 'dob_month', 'gender', 'tenure')]

You can also select a subset using the subset() function and the "select" argument:

pf_subset <- subset(pf, select = -c(userid, year_joined, year_joined_bucket))

The - sign in the "select" value indicates all but the listed columns.

You may find in your matrix that variable labels are on the outer edges of the scatterplot matrix, rather than on the diagonal. If you want labels in the diagonal, you can set the axisLabels = 'internal' argument in your ggpairs command.
***
```{r}
#install.packages('GGally')
library(GGally)
theme_set(theme_minimal(20))

# set the seed for reproducible results
set.seed(1836)
pf_subset <- pf[, c(2:15)]
names(pf_subset)
#ggpairs(pf_subset[sample.int(nrow(pf_subset), 1000), ], axisLabels = 'internal')
```

### Even More Variables
Notes:

***

### Heat Maps
Notes:
Melt data frames in R (http://www.r-bloggers.com/melt/)

Micro-array/Gene Expression Data (https://s3.amazonaws.com/udacity-hosted-downloads/ud651/nci.tsv)

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
library(reshape2)
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

