Lesson 5
========================================================

### Multivariate Data
Notes:
```{r}
getwd()
#setwd('/Users/pure/Private_Local_Data/Study/Udacity/DataAnalyst/Lessons/4.\ Exploratory\ Data\ Analysis/eda-course-materials/lesson3/')
setwd('/media/pure/SharedData/Development/Personal_Dev/DataAnalyst/Lessons/4. Exploratory Data Analysis/eda-course-materials/lesson3/')
list.files()

library(ggplot2)
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')
### You can also read in the data using the following code:
#pf < - read.delim('pseudo_facebook.tsv')
```
***

### Moira Perceived Audience Size Colored by Age
Notes:
Moira's Investigation from Lesson 3 
(https://classroom.udacity.com/courses/ud651/lessons/755618712/concepts/8140986010923)

Bernstein, M. S., Bakshy, E., Burke, M., & Karrer, B. (2013). Quantifying the invisible audience in social networks. In Proceedings of the SIGCHI Conference on Human Factors in Computing Systems (CHI 2013), pp. 21-30. 
(http://hci.stanford.edu/publications/2013/invisibleaudience/invisibleaudience.pdf)
***

### Third Qualitative Variable
Notes:

Important Notice! Please note that in newer versions of dplyr (0.3.x+), the syntax %.% has been deprecated and replaced with %>%. Videos in the course use %.%, which will produce warning messages. If you answer using the chain operator, you should use %>% instead.

Another warning: Version 0.4.0 of dplyr has a bug when using the median function on the summarize layer, depending on the nature of the data being summarized. You may need to cast the data as a numeric (float) type to get the expected results, e.g. median(as.numeric(var)).

You can include multiple variables to split the data frame when using group_by() function in the dplyr package.

new_groupings <- group_by(data, variable1, variable2)

Repeated use of summarise() and group_by(): (http://stackoverflow.com/questions/21653295/dplyr-issues-with-group-by) The summarize function will automatically remove one level of grouping (the last group it collapsed).


```{r}
library(ggplot2)
suppressMessages(library(dplyr))
```

```{r}
ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 4)
```

```{r}
ggplot(aes(x = age, y = friend_count), 
       data = subset(pf, !is.na(gender))) +
  geom_line(aes(color = gender), stat = 'summary', fun.y = median)
```

```{r Third Qualitative Variable}
pf.fc_by_age_gender <- pf %>%
  filter(!is.na(gender)) %>%
  group_by(age, gender) %>%
  summarise(mean_friend_count = mean(friend_count),
            median_friend_count = median(friend_count),
            n = n()) %>%
  ungroup() %>%
  arrange(age)
```
Write code to create a new data frame, called 'pf.fc_by_age_gender', that contains
information on each age AND gender group.

The data frame should contain the following variables:

  mean_friend_count,
  median_friend_count,
  n (the number of users in each age and gender grouping)

Here is an example of the structure of your data frame. Your
data values will be different. Note that if you are grouping by
more than one variable, you will probably need to call the
ungroup() function. 

  age gender mean_friend_count median_friend_count    n
1  13 female          247.2953                 150  207
2  13   male          184.2342                  61  265
3  14 female          329.1938                 245  834
4  14   male          157.1204                  88 1201

***

### Plotting Conditional Summaries
Notes:
Create a line graph showing the median friend count over the ages
for each gender. Be sure to use the data frame you just created,
pf.fc_by_age_gender.

Your code should look similar to the code we used to make the plot the first time. 
It will not need to make use of the stat and fun.y parameters.

ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(gender))) +
  geom_line(aes(color = gender), stat = 'summary', fun.y = median)
  
You could also use color = gender inside the aes() wrapper of ggplot().


```{r Plotting Conditional Summaries}
ggplot(aes(x = age, y = median_friend_count),
       data = pf.fc_by_age_gender) +
  geom_line(aes(color = gender))
```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:
You can also restructure the data using the tidyr package. You can review examples and how to use the package in the Data Wrangling with R (https://s3.amazonaws.com/udacity-hosted-downloads/ud651/DataWranglingWithR.pdf) pdf.

The code to change the data frame from long format to wide (or tidy format) is shown below. I encourage you to read the Data Wrangling pdf and write code using the tidyr package before looking at the solution below. 

install.packages("tidyr")
library(tidyr)

spread(subset(pf.fc_by_age_gender, 
       select = c('gender', 'age', 'median_friend_count')), 
       gender, median_friend_count)
       
I think you will find the tidyr package easier to use than the reshape2 package. Both packages can get the job done.

An Introduction to reshape2 by Sean Anderson (http://www.seananderson.ca/2013/10/19/reshape.html)

Converting Between Long and Wide Format (http://www.cookbook-r.com/Manipulating_data/Converting_data_between_wide_and_long_format/)

Melt Data Frames (http://www.r-bloggers.com/melt/)
***
```{r}
#install.packages("tidyr")
library(tidyr)
```
```{r}
spread(subset(pf.fc_by_age_gender, 
              select = c('gender', 'age', 'median_friend_count')),
       gender, median_friend_count)
```


### Reshaping Data
Notes:
It’s important to use quotes around the variable name that is assigned tovalue.var. 

We could also create a similar data frame using the dplyr and tidyr packages:

```{r reshape}
#install.packages('reshape2')
library(reshape2)

pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender,
                                  age ~ gender,
                                  value.var = 'median_friend_count')

head(pf.fc_by_age_gender.wide)
```


```{r Alternative method for reshaping}
pf.fc_by_age_gender.wide <-
  subset(pf.fc_by_age_gender[c('age', 'gender', 'median_friend_count')],
         !is.na(gender)) %>%
  spread(gender, median_friend_count) %>%
  mutate(ratio = male / female)

head(pf.fc_by_age_gender.wide)
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
ggplot(aes(x = age, y = male/female), data = pf.fc_by_age_gender.wide) +
  geom_line() +
  geom_hline(yintercept = 1, linetype = 'dashed', color = 'blue')
```

***

### Third Quantitative Variable
Notes:
Create a variable called year_joined in the pf data frame using the variable tenure and 2014 as the reference year.

The variable year_joined should contain the year that a user joined facebook.

See the Instructor Notes for three hints if you get stuck. Scroll down slowly to see one hint at a time if you would like some guidance.

A common mistake is to use tenure rather than pf$tenure or with(pf, tenure...). Remember that you need to access the variable in the data frame. This is not one of the hints! :) 

Hint 1: Divide the tenure variable by a number. Tenure is measured in days, but we want to convert it to years. 

Hint 2: Subtract tenure measured in years from 2014. What does the decimal portion represent? Should we round up or round down to the closest year? 

Hint 3: You can use the floor() function to round down to the nearest integer. You can use the ceiling() function to round up to the nearest integer. Which one should you use?

```{r Third Quantitative Variable}
pf$year_joined <- floor(2014 - pf$tenure/365)
```

***

### Cut a Variable
Notes:

Create a new variable in the data frame called year_joined.bucket by using the cut function on the variable year_joined.

You need to create the following buckets for the new variable, year_joined.bucket

      (2004, 2009]
      (2009, 2011]
      (2011, 2012]
      (2012, 2014]

Note that a parenthesis means exclude the year and a bracket means include the year.

Look up the documentation for cut or try the link in the Instructor Notes to accomplish this task.

The Cut Function (http://www.r-bloggers.com/r-function-of-the-day-cut-2/)

Be sure to close your parentheses!

A common mistake is to use year_joined rather than pf$year_joined or with(pf, year_joined...). Remember that you need to access the variable in the data frame.

```{r}
summary(pf$year_joined)
```
```{r}
table(pf$year_joined)
```


```{r Cut a Variable}
pf$year_joined.bucket <- cut(pf$year_joined, breaks = c(2004, 2009, 2011, 2012, 2014))
```

***

### Plotting it All Together
Notes:

```{r}
table(pf$year_joined.bucket, useNA = 'ifany')
```

```{r Plotting it All Together}

```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}

```

***

### Friending Rate
Notes:

```{r Friending Rate}

```

***

### Friendships Initiated
Notes:

What is the median friend rate?

What is the maximum friend rate?

```{r Friendships Initiated}

```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}

```

***

### Number of Purchases
Notes:

```{r Number of Purchases}

```

***

### Prices over Time
Notes:

```{r Prices over Time}

```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}

```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

