---
output: html_document
editor_options: 
  chunk_output_type: console
---
Lesson 6
========================================================

### Welcome
Notes:
In this lesson, Solomon will be using a linear regression model to predict diamond price using other variables in the diamonds dataset. If you are not familiar with linear regression, you may want to take a break and go through Lesson 15 of Udacity's Intro to Inferential Statistics course (https://classroom.udacity.com/courses/ud201), which covers linear regression. When you're done, you'll be ready to come back and apply your new knowledge to the diamonds dataset!
***

### Scatterplot Review
Let's start by examining two variables in the data set. The scatterplot is a powerful tool to help you understand the relationship between two continuous variables.

We can quickly see if the relationship is linear or not. In this case, we can use a variety of diamond characteristics to help us figure out whether the price advertised for any given diamond is reasonable or a rip-off.

Let's consider the price of a diamond and it's carat weight. Create a scatterplot of price (y) vs carat weight (x).

Limit the x-axis and y-axis to omit the top 1% of values.

Hint: Use the quantile() function inside of xlim and ylim to omit the top 1% of values for each variable.
```{r Scatterplot Review}
ggplot(aes(x = carat, y = price), data = subset(diamonds, price < quantile(price, probs = .99) & carat < quantile(carat, probs = .99))) +
  geom_point(fill = I('#F79420'), color = I('black'), shape = 21)
```

```{r Alternative code}
qplot(data = diamonds, x = carat, y = price,
      xlim = c(0, quantile(diamonds$carat, 0.99)),
      ylim = c(0, quantile(diamonds$price, 0.99))) +
  geom_point(fill = I('#F79420'), color = I('black'), shape = 21)
```

```{r Alternative code}
ggplot(aes(x = carat, y = price), data = diamonds) +
  scale_x_continuous(lim = c(0, quantile(diamonds$carat, 0.99))) +
  scale_y_continuous(lim = c(0, quantile(diamonds$price, 0.99))) +
  geom_point(fill = I('#F79420'), color = I('black'), shape = 21)
```


***

### Price and Carat Relationship
Response:
```{r}
ggplot(aes(x = carat, y = price), data = diamonds) +
  scale_x_continuous(lim = c(0, quantile(diamonds$carat, 0.99))) +
  scale_y_continuous(lim = c(0, quantile(diamonds$price, 0.99))) +
  geom_point(fill = I('#F79420'), color = I('black'), shape = 21, alpha = 1/4) +
  stat_smooth(method = 'lm')
```

***

### Frances Gerety
Notes:

#### A diamonds is


***

### The Rise of Diamonds
Notes:
The Counte of Monte Cristo (https://www.youtube.com/watch?v=PJyRNeCErNE) and Knocked Up (https://www.youtube.com/watch?v=Be3sZKmfb9g) are two movies that have proposals without engagement rings.

These are the only two movies that came to mind...after A LOT of time. Can you think of any others? If so, share it in the discussions!
***

### ggpairs Function
Notes:
You can click on the packages tab in RStudio to determine which packages have been installed.

You may receive a message when installing the new packages. If so, click cancel, clear your workspace, and try installing the packages again.

In this video, Solomon works with the plyr package. We worked with the dplyr package to manipulate data frames and to create new ones throughout the course. dplyr is the latest version of plyr that is specifically for working with data frames.

Similarly, we worked with the reshape2 package, which is the newest version of the reshape package.

ggpairs output (https://s3.amazonaws.com/udacity-hosted-downloads/ud651/ggpairs_landscape.pdf) When you duplicate the plot matrix on your local machine, you may want to add a axisLabels = 'internal' argument to your ggpairs function call to have the variable names on the diagonal of the matrix rather than on the outside.

```{r ggpairs Function}
# install these if necessary
install.packages('GGally') # for multiple particular plot
install.packages('scales') # for a variey of things
install.packages('memisc') # to summarize regression
install.packages('lattice') # 
install.packages('MASS') # for various functions  
install.packages('car') # to recode variables
install.packages('reshape2') # 
install.packages('plyr') # to create interesting summaries and transmissions

# load the ggplot graphics package and the others
library(ggplot2)
library(GGally)
library(scales)
library(memisc)

# sample 10,000 diamonds from the data set
set.seed(20022012)
diamond_samp <- diamonds[sample(1:length(diamonds$price), 10000), ]

ggpairs(diamond_samp, #axisLabels = 'internal',
  lower = list(continuous = wrap("points", shape = I('.'))),
  upper = list(combo = wrap("box", outlier.shape = I('.'))))
```

What are some things you notice in the ggpairs output?
Response:

***

### The Demand of Diamonds
Notes:
Create two histograms of the price variable and place them side by side on one output image.

We’ve put some code below to get you started.

The first plot should be a histogram of price and the second plot should transform the price variable using log10.

Set appropriate bin widths for each plot. ggtitle() will add a title to each histogram.

You can self-assess your work with the plots in the solution video.

```{r The Demand of Diamonds}
library(gridExtra)

plot1 <- ggplot(aes(x = price), data = diamonds) + 
  geom_histogram(binwidth = 100, fill = I('#099DD9')) +
  ggtitle('Price')

plot2 <- ggplot(aes(x = price), data = diamonds) +
  geom_histogram(binwidth = 0.01, fill = I('#F79420')) +
  scale_x_log10() +
  ggtitle('Price (log10)')

grid.arrange(plot1, plot2, ncol=2)
```

***

### Connecting Demand and Price Distributions
Notes:
When looking at these plots, what do you notice? Think specifically about the two peaks in the transformed plot and how it relates to the demand for diamonds.

Answer: 
There are two peaks in log scale plot, while in the normal plot I cannot distingush shuch peask. 
***

### Scatterplot Transformation

Basic Structure of a Function (https://www.youtube.com/watch?v=Z1wB1rHAYzQ&list=PLOU2XLYxmsIK9qQfztXeybpHvru-TrqAP)

```{r Scatterplot Transformation}
ggplot(aes(x = carat, y = price), data = diamonds) +
  geom_point() +
  scale_y_continuous(trans = log10_trans()) +
  ggtitle('Price (log10) by Carat')
```


### Create a new function to transform the carat variable

```{r cuberoot transformation}
cuberoot_trans = function() trans_new('cuberoot', transform = function(x) x^(1/3),
                                      inverse = function(x) x^3)
```

#### Use the cuberoot_trans function
```{r Use cuberoot_trans}
ggplot(aes(carat, price), data = diamonds) + 
  geom_point(alpha = 1/2, size = 0.75, position = 'jitter') + 
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat')
```

***

### Overplotting Revisited

```{r Sort and Head Tables}
head(sort(table(diamonds$carat), decreasing = T))
head(sort(table(diamonds$price), decreasing = T))
```


```{r Overplotting Revisited}
ggplot(aes(carat, price), data = diamonds) + 
  geom_point() + 
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat')
```

***

### Other Qualitative Factors
Notes:
What makes a diamond sparkle? (http://www.bluenile.com/diamonds/diamond-cut)

Does clarity affect the sparkle? (http://www.bluenile.com/diamonds/diamond-clarity)

What's a jeweler's loupe? (http://en.wikipedia.org/wiki/Loupe)

***

### Price vs. Carat and Clarity
Adjust the code below to color the points by clarity.

A layer called scale_color_brewer() has been added to adjust the legend and provide custom colors.

See if you can figure out what it does. Links to resources are in the Instructor Notes.

You will need to install the package RColorBrewer in R to get the same colors and color palettes.
```{r}
# install and load the RColorBrewer package
#install.packages('RColorBrewer', dependencies = TRUE)
library(RColorBrewer)
```

ggplot2: scale_colour_brewer
ggplot2: Color Brewer Palettes and Safe Colors/#palettes-color-brewer)
ggplot2: Legends/)

override.aes() change legend symols to be larger or more be transparent than the plot's ones. 

```{r Price vs. Carat and Clarity}
ggplot(aes(x = carat, y = price, colour = clarity), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
    guide = guide_legend(title = 'Clarity', reverse = T,
    override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
    breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
    breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Clarity')
```

***

### Clarity and Price
Response:

***

### Price vs. Carat and Cut

Alter the code below.
```{r Price vs. Carat and Cut}
ggplot(aes(x = carat, y = price, color = cut), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Clarity', reverse = T,
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and cut')
```

***

### Cut and Price
Response:

***

### Price vs. Carat and Color

Alter the code below.
```{r Price vs. Carat and Color}
ggplot(aes(x = carat, y = price, color = color), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Color', reverse = FALSE,
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Color')
```

***

### Color and Price
Response:

***

### Linear Models in R
Notes:

Response:

***

### Building the Linear Model
Notes:

```{r Building the Linear Model}
m1 <- lm(I(log(price)) ~ I(carat^(1/3)), data = diamonds)
m2 <- update(m1, ~ . + carat)
m3 <- update(m2, ~ . + cut)
m4 <- update(m3, ~ . + color)
m5 <- update(m4, ~ . + clarity)
mtable(m1, m2, m3, m4, m5)
```

Notice how adding cut to our model does not help explain much of the variance
in the price of diamonds. This fits with out exploration earlier.

***

### Model Problems
Video Notes:

Research:
(Take some time to come up with 2-4 problems for the model)
(You should 10-20 min on this)

Response:

***

### A Bigger, Better Data Set
Notes:

```{r A Bigger, Better Data Set}
install.package('bitops')
install.packages('RCurl')
library('bitops')
library('RCurl')

diamondsurl = getBinaryURL("https://raw.github.com/solomonm/diamonds-data/master/BigDiamonds.Rda")
load(rawConnection(diamondsurl))
```

The code used to obtain the data is available here:
https://github.com/solomonm/diamonds-data

## Building a Model Using the Big Diamonds Data Set
Notes:

```{r Building a Model Using the Big Diamonds Data Set}

```


***

## Predictions

Example Diamond from BlueNile:
Round 1.00 Very Good I VS1 $5,601

```{r}
#Be sure you’ve loaded the library memisc and have m5 saved as an object in your workspace.
thisDiamond = data.frame(carat = 1.00, cut = "V.Good",
                         color = "I", clarity="VS1")
modelEstimate = predict(m5, newdata = thisDiamond,
                        interval="prediction", level = .95)
```

Evaluate how well the model predicts the BlueNile diamond's price. Think about the fitted point estimate as well as the 95% CI.

***

## Final Thoughts
Notes:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

