---
title: "Load Data"
author: "Seongmuk Gang"
date: "18 December 2017"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Set Wroking Directory
setwd('/mnt/SharedData/Development/Personal_Dev/DataAnalyst/HomeWork/P4_EDA/')

## Import Libraries
```{r import library}
library(ggplot2)
library(dplyr)
```

## Features of data
I'm curious about why people take loans and what kinds of factors have effects on people to take a loan. There must be as well relationships between those factors, and it also must be hard to show clear correlations of such complex data. Therefore, I want to start Exploratory Data Analysis before I deeply dive into it.

First of all, I loaded csv data to R dataframe, and searched a length and number of variables. There are 113,937 debtors with 81 factors in the data.

```{r load csv}
pld <- read.csv('prosperLoanData.csv')
names(pld)
dim(pld)
```

It's necessray to know about types of variables, whether they are continuous or discrete
```{r}
str(pld)
```


```{r}
ggplot(data=pld,aes(LoanStatus))+geom_bar(color=I('black'),fill=I('#56B4E9'))
```

