Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:
```{r}
getwd()
setwd('/Users/pure/Private_Local_Data/Study/Udacity/DataAnalyst/Lessons/4.\ Exploratory\ Data\ Analysis/eda-course-materials/lesson3/')
list.files()

library(ggplot2)
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')
### You can also read in the data using the following code:
#pf < - read.delim('pseudo_facebook.tsv')
```
***

### Scatterplots
Notes:
http://hci.stanford.edu/publications/2013/invisibleaudience/invisibleaudience.pdf

```{r Scatterplots}
qplot(x = age, y = friend_count, data = pf)
qplot(age, friend_count, data = pf)
```

The equivalent ggplot syntax for the scatterplot:
```{r}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point()
```

***

#### What are some things that you notice right away?
Response:

***

### ggplot Syntax
Notes:
ggplot2 geoms: http://docs.ggplot2.org/current/
ggplot2 tutorial: http://bbs.ceb-institute.org/wp-content/uploads/2011/09/handout_ggplot2.pdf
```{r ggplot Syntax}
ggplot(aes(x = age, y = friend_count), data = pf)+
  geom_point() +
  xlim(13, 90)
```
```{r ggplot Syntax}
ggplot(aes(x = age, y = friend_count), data = pf)+
  geom_jitter() +
  xlim(13, 90)
```

***

### Overplotting
Notes:
Overplotting makes it difficult to tell how many points are in each region.

```{r Overplotting}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_jitter(alpha = 1/20) +
  xlim(13, 90)
```

#### What do you notice in the plot?
Response:

***

### Coord_trans()
Notes:
http://docs.ggplot2.org/current/coord_trans.html

With this plot, it's much easier to see the distribution of friend count, conditional, and age.
```{r Coord_trans()}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point(alpha = 1/20) +
  xlim(13, 90)
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point(alpha = 1/20, position = position_jitter(h = 0)) +
  xlim(13, 90) +
  coord_trans(y = 'sqrt')
```

#### What do you notice?

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}
ggplot(aes(x = age, y = friendships_initiated), data = pf) +
  geom_point(alpha = 1/20, position = position_jitter(h = 0)) +
  xlim(13, 90) +
  coord_trans(y = 'sqrt')
```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:
Important Notice! Please note that in newer versions of dplyr (0.3.x+), the syntax %.% has been deprecated and replaced with %>%.

Another warning: Version 0.4.0 of dplyr has a bug when using the median function on the summarize layer, depending on the nature of the data being summarized. You may need to cast the data as a numeric (float) type when using it on your local machine, e.g. median(as.numeric(var)).

Learn more about the dplyr package (http://blog.rstudio.org/2014/01/17/introducing-dplyr/). 

Introduction to dplyr (http://rstudio-pubs-static.s3.amazonaws.com/11068_8bc42d6df61341b2bed45e9a9a3bf9f4.html)

The following tutorials are presented by Hadley Wickham at useR 2014.

- Introduction of dplyr (http://www.r-bloggers.com/hadley-wickham-presents-dplyr-at-user-2014/)
- dplyr Tutorial Part 1 (http://www.r-bloggers.com/hadley-wickhams-dplyr-tutorial-at-user-2014-part-1/)
- dplyr Tutorial Part 2 
(http://www.r-bloggers.com/hadley-wickhams-dplyr-tutorial-at-user-2014-part-2/)

There are other ways to work with data and create new data frames without using the dplyr package. Learn about the R functions lapply, tapply, and split in a blog post (http://rollingyours.wordpress.com/2014/10/20/the-lapply-command-101/).

For more on geom_line(), you can check the documentation here (http://docs.ggplot2.org/current/geom_path.html).

```{r Conditional Means}
#install.packages('dplyr')
#library(dplyr)

## filter()
## group_by()
## mutate()
## arrange()

age_groups <- group_by(pf, age)
pf.fc_by_age <- summarise(age_groups,
                          friend_count_mean = mean(friend_count),
                          friend_count_median = median(friend_count),
                          n = n())
head(pf.fc_by_age)

pf.fc_by_age <- arrange(pf.fc_by_age, age) # ordering
head(pf.fc_by_age)
```

## Conditional Means Alternate Code
```{r}
pf.fc_by_age <- pf %>%
  group_by(age) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %>%
  arrange(age)

head(pf.fc_by_age)
```

Create your plot!

```{r Conditional Means Plot}
ggplot(aes(x = age, y = friend_count_mean), data = pf.fc_by_age) +
  geom_line()
```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x = age, y = friend_count), data = pf) +
  xlim(13, 90) +
  geom_point(alpha = 0.05, 
             position = position_jitter(h = 0), 
             color = 'orange') +
  coord_trans(y = 'sqrt') +
  geom_line(stat = 'summary', fun.y = mean) +
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = .1),
            linetype = 2, color = 'blue') +
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = .5),
            color = 'blue') +
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = 0.9), 
            linetype = 2, color = 'blue')
```

#### What are some of your observations of the plot?
Response:
```{r}
ggplot(aes(x = age, y = friend_count), data = pf) +
  coord_cartesian(xlim = c(13, 70), ylim = c(0, 1000)) +
  geom_point(alpha = 0.05, 
             position = position_jitter(h = 0), 
             color = 'orange') +
  geom_line(stat = 'summary', fun.y = mean) +
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = .1),
            linetype = 2, color = 'blue') +
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = .5),
            color = 'blue') +
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = 0.9), 
            linetype = 2, color = 'blue')
```

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:
A Visual Guide to Correlation (https://s3.amazonaws.com/udacity-hosted-downloads/ud651/correlation_images.jpeg)

Correlation Coefficient (http://www.r-tutor.com/elementary-statistics/numerical-measures/correlation-coefficient)

Intro to Inferential Statistics- Correlation (https://classroom.udacity.com/courses/ud201/lessons/1345848540/concepts/1715827370923)

Correlation coefficients are often denoted with the greek letter ρ (rho), in addition to the letter r.

The default method for computing the correlation coefficient is Pearson, and this is true for most statistical software. You do not need to pass the method parameter when calculating the Pearson Product Moment Correlation.
```{r Correlation}
cor(pf$age, pf$friend_count, method = 'pearson')
```

```{r alternative Correlation}
with(pf, cor.test(age, friend_count, method = 'pearson'))
```
Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:
What are monotonic functions (http://en.wikipedia.org/wiki/Monotonic_function)?

Definition: 
ostensibly - apparently or purportedly, but perhaps not actually

Use the subset command or bracket notation for this exercise. The grader will not accept other techniques.

```{r Correlation on Subsets}
with(subset(pf, age<=70), cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:
Correlation Methods: Pearson's r, Spearman's rho, and Kendall's tau (http://www.statisticssolutions.com/correlation-pearson-kendall-spearman/)
***

## Create Scatterplots
Notes:

```{r}
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) +
  geom_point()
```

***

### Strong Correlations
Notes:
The correlation coefficient is invariant under a linear transformation of either X or Y, and the slope of the regression line when both X and Y have been transformed to z-scores is the correlation coefficient.

It's important to note that we may not always be interested in the bulk of the data. Sometimes, the outliers ARE of interest, and it's important that we understand their values and why they appear in the data set.
```{r Strong Correlations}
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) +
  geom_point() +
  xlim(0, quantile(pf$www_likes_received, 0.95)) +
  ylim(0, quantile(pf$likes_received, 0.95)) +
  geom_smooth(method = 'lm', color = 'red')
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
cor.test(pf$likes_received, pf$www_likes_received)
```

Response:

***

### Moira on Correlation
Notes:
For regression, variables sould be independent.
***

### More Caution with Correlation
Notes:
Argument matching (when not providing them by name) in R is a bit complex.

First, arguments (or parameters) can be matched by name. If a parameter matches exactly, it is "removed" from the argument list and the remaining unnamed arguments are matched in the order that they are listed in the function definition.

R does the following to match arguments... 

checks for exact match of named argument
checks for a partial match of the argument
checks for a positional match
If R does not find a match for a parameter, it typically throws an "unused" parameter error.

Type str(functionName) to find the order of the parameters and learn more about the parameters of an R function. 

The example covered in the next few videos comes from a practice problem in Weisberg, S. (2005). Applied Linear Regression, 3rd edition. New York: Wiley (http://users.stat.umn.edu/~sandy/alr3ed/website/).

```{r More Caution With Correlation}
# install.packages('alr3')
# library(alr3)

data("Mitchell")
?Mitchell
```

Create your plot!

```{r Temp vs Month}
ggplot(aes(x = Month, y = Temp), data = Mitchell) +
  geom_point()
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
cor.test(Mitchell$Month, Mitchell$Temp)
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
range(Mitchell$Month)

ggplot(data = Mitchell, aes(x = Month, y = Temp)) +
  geom_point() +
  scale_x_continuous(breaks = seq(0, 203, 12))
```

***

### A New Perspective
You could also get perspective on this data by overlaying each year's data on top of each other, giving a clear, generally sinusoidal graph. You can do this by using the R's modulus (http://en.wikipedia.org/wiki/Modular_arithmetic) operator %% in your code. Try running the code below!
```{r}
ggplot(aes(x=(Month%%12),y=Temp), data=Mitchell)+
  geom_point()
```

Data Visualization Pioneers

John Tukey (http://en.wikipedia.org/wiki/John_Tukey)

William Playfair (http://en.wikipedia.org/wiki/William_Playfair)

William Playfair and the Psychology of Graphs (http://www.psych.utoronto.ca/users/spence/Spence%20(2006).pdf)

There are other measures of associations that can detect this. The dcor.ttest() function in the energy package implements a non-parametric test of the independence of two variables. While the Mitchell soil dataset is too coarse to identify a significant dependency between "Month" and "Temp", we can see the difference between dcor.ttest and cor.test through other examples, like the following:

```{r}
#install.packages(energy)
library(energy)

x <- seq(0, 4*pi, pi/20)
y <- cos(x)
qplot(x = x, y = y)
dcor.ttest(x, y)
```

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:
Assume the reference date for calculating age is December 31, 2013 and that the age variable gives age in years at the end of 2013.

The variable age_with_months in the data frame pf should be a decimal value. For example, the value of age_with_months for a 33 year old person born in March would be 33.75 since there are nine months from March to the end of the year. Similarly, someone who was 6 years and 6 months old would have an age_with_months value of 6.5.

```{r}
ggplot(aes(x = age, y = friend_count_mean), data = pf.fc_by_age) +
  geom_line()
```
```{r}
head(pf.fc_by_age, 10)
```
```{r}
pf.fc_by_age[18:19, ]
```

```{r Understanding Noise: Age to Age Months}
pf$age_with_months <- pf$age + (1 - pf$dob_month/12)

## alternate solution:
# pf$age_with_months <- with(pf, age + (1 - dob_month / 12))
```

***

### Age with Months Means
Important Notice! Please note that in newer versions of dplyr (0.3.x+), the syntax %.% has been deprecated and replaced with %>%. Videos in the course use %.%, which will produce warning messages. If you answer using the chain operator, you should use %>% instead. Another warning: Version 0.4.0 of dplyr has a bug when using the median function on the summarize layer, depending on the nature of the data being summarized. You may need to cast the data as a numeric (float) type to get the expected results, e.g. median(as.numeric(var)).

A few additional hints follow below: 

Hint 1: Use the group_by(), summarise(), and arrange() functions in the dplyr package to split the data frame by age_with_month. Make sure you arrange by the correct variable (it's not age anymore). 

Hint 2: The code should look similar to the code when we split the data frame by age and found summaries:

age_groups <- group_by(pf, age)
pf.fc_by_age <- summarise(age_groups,
                          friend_count_mean = mean(friend_count),
                          friend_count_median = median(friend_count),
                          n = n())
pf.fc_by_age <- arrange(pf.fc_by_age, age)
head(pf.fc_by_age)

```{r Age with Months Means}
library(dplyr)

pf.fc_by_age_months <- pf %>%
  group_by(age_with_months) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(as.numeric(friend_count)),
            n =n()) %>%
  arrange(age_with_months)

head(pf.fc_by_age_months, 5)
```

```{r Alternative code}
age_with_months_groups <- group_by(pf, age_with_months)
pf.fc_by_age_months2 <- summarise(age_with_months_groups,
                                  friend_count_mean = mean(friend_count),
                                  friend_count_median = median(friend_count),
                                  n = n())
pf.fc_by_age_months2 <- arrange(pf.fc_by_age_months2, age_with_months)

head(pf.fc_by_age_months2)
```


Programming Assignment
```{r Programming Assignment}

```

***

### Noise in Conditional Means
Create a new line plot showing friend_count_mean versus the new variable,
age_with_months. Be sure to use the correct data frame (the one you created
in the last exercise) AND subset the data to investigate users with ages less
than 71.
```{r Noise in Conditional Means}
ggplot(aes(x = age_with_months, y = friend_count_mean), 
       data = subset(pf.fc_by_age_months, age_with_months < 71)) +
    geom_line()
```

***

### Smoothing Conditional Means
Notes:
Local Regression (LOESS) (http://simplystatistics.org/2014/02/13/loess-explained-in-a-gif/) explained visually on the Simply Statistics (http://simplystatistics.org/) blog.

The Details of Loess and Lowess (http://en.wikipedia.org/wiki/Local_regression)

```{r Smoothing Conditional Means}
p1 <- ggplot(aes(x = age, y = friend_count_mean),
       data = subset(pf.fc_by_age, age < 71)) +
  geom_line() +
  geom_smooth()

p2 <- ggplot(aes(x = age_with_months, y = friend_count_mean),
       data = subset(pf.fc_by_age_months, age_with_months < 71)) +
  geom_line() +
  geom_smooth()

p3 <- ggplot(aes(x = round(age / 5) * 5, y = friend_count),
             data = subset(pf, age < 71)) +
  geom_line(stat = 'summary', fun.y = mean)
library(gridExtra)
grid.arrange(p2, p1, p3, ncol = 1)
```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:
A deep dive into Bivariate Data Analysis (http://dept.stat.lsa.umich.edu/~kshedden/Courses/Stat401/Notes/401-bivariate-slides.pdf).
***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

